---
title: "Almond Profit - Informal Sensitivity Analysis"
author: "Ben Versteeg, Hope Hahn"
output: html_document
date: "2024-04-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(purrr)
library(ggpubr)
```

Tasks:
-- Develop a profit model for your almond yield (you can make this up; think about what the parameters would be) 
- you might assume a baseline profit and then adjust according to the anomaly 
- there are many ways to combine the almond yield and profit functions;you can have the profit function “call”/use the almond  yield function; or create a wrapper function that calls them in sequence (first the almond yield and then the profit function) 

-- Do a simple informal sensitivity analysis of almond yield profit using at least 2 parameters

-- Create 2 graphs of the results - you can decide what is the most meaningful graph

-- Write a short paragraph (in the Rmarkdown document) to summarize you interpretation of your model results (e.g what do ‘take away’ from your model and sensitivity analysis


## Call the almond profit function

```{r}
# Load the R script containing the almond_profit_fun function
source("almond_profit.R")
```

## Read in and prepare the data

```{r}
# Read in the CSV file and clean the data
jan_alm <- read_csv(here::here("data/clim.txt")) %>%

  # Split the single column into multiple columns based on quotes
  separate(col = "day month year wy tmax_c tmin_c precip",
           into = c("day", "month", "year", "wy", "tmax_c", "tmin_c", "precip"),
           sep = " ") %>%

  # Convert columns to numeric
  mutate(year = as.numeric(year),
         tmin_c = as.numeric(tmin_c),
         precip = as.numeric(precip)) %>%

  # Group the data by year and month
  group_by(year, month) %>%

  # Keep only the rows where month equals 1 (January)
  filter(month == 1) %>%

  # Compute the total precipitation for each year in January
  summarise(precip_sum = sum(precip, na.rm = TRUE))

```

```{r}
# Read the CSV file 
feb_alm <- read_csv(here::here("data/clim.txt")) %>%

  # Split the single column into multiple columns based on quotes
  separate(col = "day month year wy tmax_c tmin_c precip",
           into = c("day", "month", "year", "wy", "tmax_c", "tmin_c", "precip"),
           sep = " ") %>%

  # Convert columns to numeric 
  mutate(year = as.numeric(year),
         tmin_c = as.numeric(tmin_c),
         precip = as.numeric(precip)) %>%

  # Group the data by year and month
  group_by(year, month) %>%

  # Keep only the rows where month equals 2 (February)
  filter(month == 2) %>%

  # Compute the minimum temperature for each year in February
  summarise(tempc_min = min(tmin_c, na.rm = TRUE))

```
## Running the Almond profit model with default values.

```{r}
# Calculate almond profit by passing February minimum temperatures and January precipitation sums as arguments to the almond_profit_fun function
almond_profit_output <- almond_profit_fun(min_temp = feb_alm$tempc_min, precip = jan_alm$precip_sum)

paste0("Profit = ", almond_profit_output)

```

## Simple informal sensitivity analysis of almond yield profit using at least 2 parameters

```{r}
# Define ranges for acre and price_per_pound
acre_range <- seq(20, 40, by = 2)  # Create a sequence of acre values 
price_per_pound_range <- seq(1, 3, by = 0.25)  # Create a sequence of price per pound values 

# Use expand.grid to create all possible combinations of acre, price_per_pound, and average_yield
results <- expand.grid(acre = acre_range, price_per_pound = price_per_pound_range)

results$profit <- NA  # Initialize a new column 'profit' in the results data frame with NA (missing values)

# Loop over each row in the results data frame to calculate profits for each combination of parameters
for (i in 1:nrow(results)) {
  # Call the almond_profit_fun function for each combination of parameters.
  profit <- almond_profit_fun(min_temp = feb_alm$tempc_min, precip = jan_alm$precip_sum,
                              acre = results[i, "acre"], price_per_pound = results[i, "price_per_pound"])
  # Store the computed profit in the corresponding row of the 'profit' column
  results[i, "profit"] <- profit
}

```

## Create 2 graphs of the results

### First Graph
```{r}
# Visualize the results
ggplot(results, aes(x = acre, y = price_per_pound, fill = profit)) +
  geom_tile() +
  scale_fill_gradient(low = "yellow", high = "red") +
  labs(x = "Acre", y = "Price per Pound (USD)", fill = "Profit (USD)")
```

### Second Graph

```{r}
ggplot(results, aes(x = price_per_pound, y = profit, group = price_per_pound)) +
  geom_boxplot() +
  labs(title = "Profit by Price per Pound",
       y = "Profit (USD)", 
       x = "Price per Pound (USD)") +
  theme_classic()

ggplot(results, aes(x = acre, y = profit, group = acre)) +
  geom_boxplot() +
  labs(title = "Profit by Acre",
       y = "Profit (USD)", 
       x = "Acre") +
  theme_classic()
```

## Short paragraph

Based on the graphs, it appears that the profit (output) is pretty sensitive to the acre and price per pound parameters. As shown in the heatmap, Acre and Price per pound are correlated and as both increase, the profit increases as well. The 



